%!TEX root = haiku.tex
\book{\LARGE \FK 杂选}

\newpage

\chapter{\FK 一}
\setcounter{haikucounter}{0}

\begin{haiku}
    {\FH 春の風草深くても\ruby[g]{古郷}{こきょう}\ruby[g]{也}{なり}}\hfill{\FH 一茶}

    {\FK 春风呵，虽然草长得深，还是故乡呵。}
\end{haiku}

\begin{haiku}
    {\FH 世の中は地獄の上の花見哉}\hfill{\FH 一茶}

    {\FK 我们在世上，边看繁花，边朝地狱行去。}
\end{haiku}

\begin{haiku}
    {\FH ままっ子や涼み仕事にわらたたき}\hfill{\FH 一茶}

    {\FK 继子呵，乘凉时候的执事是敲稻草。}
\end{haiku}

\begin{haiku}
    {\FH 正月やえたの玄関も梅の花}\hfill{\FH 一茶}

    {\FK 秽多玄关外，正月梅花开。}
\end{haiku}

\begin{haiku}
    {\FH 門の木も\ruby[g]{先}{まづ}つつがなし\ruby[g]{夕涼}{ゆうすずみ}}\hfill{\FH 一茶}

    {\FK 门前树依然无恙，好纳晚凉。}
\end{haiku}

\begin{haiku}
    {\FH わが春やたどん一つにい\ruby[g]{小菜}{こな}\ruby[g]{一把}{いちわ}}\hfill{\FH 一茶}

    {\FK 一把小菜一块煤，此乃我春天。}
\end{haiku}

\begin{haiku}
    {\FH \ruby[g]{一抱}{ひとかか}えあれど柳は柳かな}\hfill{\FH 千代女}

    {\FK 虽有一抱之粗，但柳树仍是柳树。}
\end{haiku}

\begin{haiku}
    {\FH 音\ruby[g]{添}{そ}ふて雨にしづまる\ruby[g]{碪}{きぬた}かな}\hfill{\FH 千代女}

    {\FK 秋夜里的捶衣板声，消失在雨声中。}
\end{haiku}

\begin{haiku}
    {\FH \ruby[g]{百生}{ひゃくなり}や\ruby[g]{蔓}{つる}\ruby[g]{一筋}{ひとすじ}の心より}\hfill{\FH 千代女}

    {\FK 百瓜一藤，百念皆由心生。}
\end{haiku}

\begin{haiku}
    {\FH 筆取て千艸の花におくるゝな}\hfill{\FH 闌更}

    {\FK 援笔默逊百花艳。}
\end{haiku}

\begin{haiku}
    {\FH \ruby[g]{枯蘆}{かれあし}の日に日に折れて流れけり}\hfill{\FH 闌更}

    {\FK 枯芦苇，日日折断随流去。}
\end{haiku}

\begin{haiku}
    {\FH 翡翠の影こんこんと\ruby[g]{遡}{さかのぼ}り}\hfill{\FH 茅舎}

    {\FK 翡翠鸟影，滚滚溯流上。}
\end{haiku}

\begin{haiku}
    {\FH 父が待ちし我が待ちし朴咲きにけり}\hfill{\FH 茅舎}

    {\FK 父亲等待我等待，朴树花盛开。}
\end{haiku}

\begin{haiku}
    {\FH \ruby[g]{朴散華}{ほおさんげ}\ruby[g]{即}{すなわ}ちしれぬ\ruby[g]{行方}{ゆくへ}かな}\hfill{\FH 茅舎}

    {\FK 朴树散花，不知去向。}
\end{haiku}

\begin{haiku}
    {\FH さびしさや\ruby[g]{一尺}{いちしゃく}消えてゆく蛍}\hfill{\FH 北枝}

    {\FK 流萤断续光，一明一灭一尺间，寂寞何以堪。}
\end{haiku}

\begin{haiku}
    {\FH ちぎりきな\ruby[g]{互}{かた}みに\ruby[g]{渋}{しぶ}き柿二つ}\hfill{\FH 大江丸}

    {\FK 两个涩柿子，互相摘一个。}
\end{haiku}

\begin{haiku}
    {\FH 枯野哉つばなの時の女櫛}\hfill{\FH 西鶴}

    {\FK 荒野草木枯，何人来撷茅草花，遗落小木梳。}
\end{haiku}

\begin{haiku}
    {\FH もりもり盛りあがる雲へあゆむ}\hfill{\FH 山頭火}

    {\FK 风起云涌，云涌云涌，步上白云。}
\end{haiku}

\begin{haiku}
    {\FH ひとり山越えてまた山}\hfill{\FH 山頭火}

    {\FK 不断地翻越，不断地翻越，依旧是青山。}
\end{haiku}

\begin{haiku}
    {\FH しみじみ食べる飯ばかりの飯である}\hfill{\FH 山頭火}

    {\FK 细细品味只有饭的饭。}
\end{haiku}

\begin{haiku}
    {\FH けふも\ruby[g]{托鉢}{たくはつ}こゝもかしこも花ざかり}\hfill{\FH 山頭火}

    {\FK 日日托钵，处处百花盛开。}
\end{haiku}

\begin{haiku}
    {\FH 親馬仔馬みんな戻ってくるあたゝかし}\hfill{\FH 山頭火}

    {\FK 大马小马，温暖齐归。}
\end{haiku}

\begin{haiku}
    {\FH 行く\ruby[g]{我}{われ}もにほへ花野を来る一人}\hfill{\FH 言水}

    {\FK 一人打自花野来，我亦染香花野去。}
\end{haiku}

\begin{haiku}
    {\FH 桜散る日さへ夕と成にけり}\hfill{\FH 樗良}

    {\FK 樱花飘散，朝开夕落。}
\end{haiku}

\begin{haiku}
    {\FH 子に\ruby[g]{飽}{あ}くと申す人には花もなし}\hfill{\FH 芭蕉}

    {\FK 有人不爱子，花不为伊开。}
\end{haiku}

\begin{haiku}
    {\FH 目には青葉山ほととぎす\ruby[g]{初鰹}{はつがつお}}\hfill{\FH 素堂}

    {\FK 青山绿叶杜鹃啼，更有鲣鱼初上市。}
\end{haiku}

\begin{haiku}
    {\FH かげろふやほろほろ落つる岸の砂}\hfill{\FH 土芳}

    {\FK 阳炎袅袅升，干沙簌簌落。}
\end{haiku}

\begin{haiku}
    {\FH 大原や蝶の出て舞う朧月}\hfill{\FH 丈草}

    {\FK 大原月朦胧，孤蝶翩来飞。}
\end{haiku}

\begin{haiku}
    {\FH 草霞み水に声なき日ぐれ哉}\hfill{\FH 蕪村}

    {\FK 晚霞草迷蒙，流水自无声。}
\end{haiku}

\begin{haiku}
    {\FH 風おもく人甘くなりて春くれぬ}\hfill{\FH 暁台}

    {\FK 风重人娇春将归。}
\end{haiku}

\begin{haiku}
    {\FH 萩芒来年\ruby{逢}{あは}んさりながら}\hfill{\FH 子規}

    {\FK 胡枝子和芒草，明年也想见到啊，可是啊可是。\footnote{\FT 这是子规于漱石去英国留学前，于病榻上为他写下的句子。漱石返日后，子规已病故。​​​}}
\end{haiku}

\begin{haiku}
    {\FH 春雨のわれまぼろしに近き身ぞ}\hfill{\FH 子規}

    {\FK 春雨细蒙蒙，我身近幻影。}
\end{haiku}

\begin{haiku}
    {\FH 鶯の脛の寒さよ竹の中}\hfill{\FH 紅葉}

    {\FK 竹里黄莺足胫寒。}
\end{haiku}

\begin{haiku}
    {\FH 蝶々のもの食ふ音の静かさよ}\hfill{\FH 虚子}

    {\FK 蝴蝶啮物声悄静。}
\end{haiku}

\begin{haiku}
    {\FH 日盛りに蝶のふれ合ふ音すなり}\hfill{\FH 青青}

    {\FK 盛夏阳光里，听见蝴蝶相触声。}
\end{haiku}

\begin{haiku}
    {\FH 死火山の膚つめたくて草いちご}\hfill{\FH 蛇笏}

    {\FK 死火山，冰冷山肌缀草莓。}
\end{haiku}

\begin{haiku}
    {\FH ふりかねてこよひになりぬ月の雨}\hfill{\FH 尚白}

    {\FK 忍到今宵偏月雨。}
\end{haiku}

\begin{haiku}
    {\FH 魚くふて口なまぐさし昼の雪}\hfill{\FH 成美}

    {\FK 吃鱼嘴腥味，闲看午后雪。}
\end{haiku}

\begin{haiku}
    {\FH 月さすや谷をさまよふ蛍どち}\hfill{\FH 石鼎}

    {\FK 山谷明月光，流萤皆彷徨。}
\end{haiku}

\begin{haiku}
    {\FH 花影婆娑と踏むべくありぬ岨の月}\hfill{\FH 石鼎}

    {\FK 花影婆娑欲踏踩，悬崖樱树月色明。}
\end{haiku}

\begin{haiku}
    {\FH 木より木に通へる風の春浅き}\hfill{\FH 亜浪}

    {\FK 风穿树间春初浅。}
\end{haiku}

\begin{haiku}
    {\FH ふりむけば灯とぼす関や夕霞}\hfill{\FH 太祇}

    {\FK 过得城关后，回收初看灯火瘦，天际晚霞秀。}
\end{haiku}

\begin{haiku}
    {\FH \ruby[g]{十団子}{とおだご}も\ruby[g]{小粒}{こつぶ}になりぬ秋の風}\hfill{\FH 許六}

    {\FK 萧瑟秋风起，江米团子亦变小。}
\end{haiku}

\begin{haiku}
    {\FH 蝶老いてたましひ菊にあそぶ哉}\hfill{\FH 星布女}

    {\FK 可怜蝶已老，游魂戏晚菊。}
\end{haiku}

\chapter{\FK 二}
\setcounter{haikucounter}{0}

\begin{haiku}
    {\FH あさがほに我は飯くふおとこ哉}\hfill{\FH 芭蕉}

    \vin{} Gazing at the flowers
    \vin{} \vin{} Of the morning glory,
    \vin{} \vin{} \vin{} I eat my breakfast.
\end{haiku}

\begin{haiku}
    {\FH いざ子供走りありかん玉霰}\hfill{\FH 芭蕉}

    \vin{} Look, children,
    \vin{} \vin{} Hail-stones!
    \vin{} \vin{} \vin{} Let's rush out!
\end{haiku}

\begin{haiku}
    {\FH いざ行かん雪見にころぶ所まで}\hfill{\FH 芭蕉}

    \vin{} Now let's be off!
    \vin{} \vin{} Let's go snow-viewing till
    \vin{} \vin{} \vin{} We tumble down!
\end{haiku}

\begin{haiku}
    {\FH きみ火をたけよき物見せん雪丸げ}\hfill{\FH 芭蕉}

    \vin{} You light the fire
    \vin{} \vin{} And I'll show you something nice ---
    \vin{} \vin{} \vin{} A ball of snow!
\end{haiku}

\begin{haiku}
    {\FH さびしさを問てくれぬか桐一葉}\hfill{\FH 芭蕉}

    \vin{} A paulonia leaf has fallen:
    \vin{} \vin{} Will you not come to me
    \vin{} \vin{} \vin{} In my loneliness?
\end{haiku}

\begin{haiku}
    {\FH ほととぎす大竹藪を漏る月夜}\hfill{\FH 芭蕉}

    \vin{} A cuckoo cried!
    \vin{} \vin{} The moon filters through
    \vin{} \vin{} \vin{} The vast bamboo grove.
\end{haiku}

\begin{haiku}
    {\FH ほととぎす鳴くや五尺のあやめ草}\hfill{\FH 芭蕉}

    \vin{} The voice of the cuckoo!
    \vin{} \vin{} And the tall
    \vin{} \vin{} \vin{} Irises.
\end{haiku}

\begin{haiku}
    {\FH 五月雨にかくれぬものや瀬田の橋}\hfill{\FH 芭蕉}

    \vin{} The summer rains:
    \vin{} \vin{} All things hidden
    \vin{} \vin{} \vin{} But the long bridge of Seta.
\end{haiku}

\begin{haiku}
    {\FH 何の木の花とは知らず匂ひ哉}\hfill{\FH 芭蕉}

    \vin{} The fragrance of some
    \vin{} \vin{} Unknown blossoming tree
    \vin{} \vin{} \vin{} Filled all my soul!
\end{haiku}

\begin{haiku}
    {\FH 名月や座に美しき顔もなし}\hfill{\FH 芭蕉}

    \vin{} The harvest moon!
    \vin{} \vin{} Among us, none
    \vin{} \vin{} \vin{} Has a face of beauty.
\end{haiku}

\begin{haiku}
    {\FH 命二つの中に生たる櫻かな}\hfill{\FH 芭蕉}

    \vin{} Our two lives:
    \vin{} \vin{} Between them is the life
    \vin{} \vin{} \vin{} Of the cherry flowers.
\end{haiku}

\begin{haiku}
    {\FH 手鼻かむ音さへ梅の匂ひ哉}\hfill{\FH 芭蕉}

    \vin{} The sound of someone
    \vin{} \vin{} Blowing his nose with his hand;
    \vin{} \vin{} \vin{} The scent of the plum flowers!
\end{haiku}

\begin{haiku}
    {\FH 日頃にくき烏も雪のあしたかな}\hfill{\FH 芭蕉}

    \vin{} How beautiful
    \vin{} \vin{} The usually hateful crow,
    \vin{} \vin{} \vin{} This snowy morning!
\end{haiku}

\begin{haiku}
    {\FH 晝見れば首筋赤き螢かな}\hfill{\FH 芭蕉}

    \vin{} By day-light
    \vin{} \vin{} The firefly has
    \vin{} \vin{} \vin{} A neck of red.
\end{haiku}

\begin{haiku}
    {\FH 朝顔やこれもまた我友ならず}\hfill{\FH 芭蕉}

    \vin{} The morning glory too
    \vin{} \vin{} Can never be
    \vin{} \vin{} \vin{} My friend.
\end{haiku}

\begin{haiku}
    {\FH 松茸やしらぬ木の葉のへばりつく}\hfill{\FH 芭蕉}

    \vin{} On the mushroom
    \vin{} \vin{} Is stuck the leaf
    \vin{} \vin{} \vin{} Of some unknown tree.
\end{haiku}

\begin{haiku}
    {\FH 棧や命をからむ蔦かつら}\hfill{\FH 芭蕉}

    \vin{} Ah, hanging bridge!
    \vin{} \vin{} Ivy-ropes
    \vin{} \vin{} \vin{} Entwine existence!
\end{haiku}

\begin{haiku}
    {\FH 石山の石にたばしるあられかな}\hfill{\FH 芭蕉}

    \vin{} The hail-stones
    \vin{} \vin{} Glance off the rocks
    \vin{} \vin{} \vin{} Of the Stony Mountain.
\end{haiku}

\begin{haiku}
    {\FH 稲妻に悟らぬ人の貴さよ}\hfill{\FH 芭蕉}

    \vin{} How admirable,
    \vin{} \vin{} He who thinks not, ``Life is fleeting''
    \vin{} \vin{} \vin{} When he sees the lightning-flash!
\end{haiku}

\begin{haiku}
    {\FH 聲にみな鳴きしまふてや蝉の殻}\hfill{\FH 芭蕉}

    \vin{} The shell of a cicada:
    \vin{} \vin{} It sang itself
    \vin{} \vin{} \vin{} Utterly away.
\end{haiku}

\begin{haiku}
    {\FH 葱白く洗ひたてたるさむさ哉}\hfill{\FH 芭蕉}

    \vin{} Just washed,
    \vin{} \vin{} How chill
    \vin{} \vin{} \vin{} The white leeks!
\end{haiku}

\begin{haiku}
    {\FH 蛇くふと聞けばおそろし雉子の聲}\hfill{\FH 芭蕉}

    \vin{} They say the pheasant
    \vin{} \vin{} Eats the snake;
    \vin{} \vin{} \vin{} How fearful now, its voice!
\end{haiku}

\begin{haiku}
    {\FH 馬ほくほく我を繪に見る夏野哉}\hfill{\FH 芭蕉}

    \vin{} The cob ambles slowly
    \vin{} \vin{} Across the summer moor;
    \vin{} \vin{} \vin{} I find myself in a picture.
\end{haiku}

\begin{haiku}
    {\FH 馬をさへ眺むる雪の朝かな}\hfill{\FH 芭蕉}

    \vin{} We gaze
    \vin{} \vin{} Even at horses,
    \vin{} \vin{} \vin{} This morn of snow!
\end{haiku}

\begin{haiku}
    {\FH 魂祭りけふも焼場のけぶりかな}\hfill{\FH 芭蕉}

    \vin{} The Festival of the Dead:
    \vin{} \vin{} But from the Burning-ground arises smoke,
    \vin{} \vin{} \vin{} Even today.
\end{haiku}

\begin{haiku}
    {\FH 鶯や餅に糞する椽の先}\hfill{\FH 芭蕉}

    \vin{} Look! the dried rice cakes
    \vin{} \vin{} At the end of the verandah ---
    \vin{} \vin{} \vin{} The uguisu is pooping on them!
\end{haiku}

\begin{haiku}
    {\FH 大名を馬からおろす櫻かな}\hfill{\FH 一茶}

    \vin{} The cherry blossoms!
    \vin{} \vin{} They have made a daimy\^o
    \vin{} \vin{} \vin{} Dismount from his horse.
\end{haiku}

\begin{haiku}
    {\FH 山水に米を搗かせて晝寝かな}\hfill{\FH 一茶}

    \vin{} I take a nap,
    \vin{} \vin{} Making the mountain water
    \vin{} \vin{} \vin{} Pound the rice.
\end{haiku}

\begin{haiku}
    {\FH 年玉や懐の子も手ゝをして}\hfill{\FH 一茶}

    \vin{} New Year's Presents:
    \vin{} \vin{} The baby also
    \vin{} \vin{} \vin{} Holds out its tiny hands.
\end{haiku}

\begin{haiku}
    {\FH 我味の柘榴に這す虱かな}\hfill{\FH 一茶}

    \vin{} Poor louse! I made it creep
    \vin{} \vin{} Upon the pomegranate that tastes
    \vin{} \vin{} \vin{} Of my sweet flesh.
\end{haiku}

\begin{haiku}
    {\FH 正月の子供に成って見たきかな}\hfill{\FH 一茶}

    \vin{} Ah! to be
    \vin{} \vin{} A child ---
    \vin{} \vin{} \vin{} On New Year's Day!
\end{haiku}

\begin{haiku}
    {\FH 狭くともいざ飛習へ庵の蚤}\hfill{\FH 一茶}

    \vin{} I'm sorry my house is so small,
    \vin{} \vin{} But practise your jumping,
    \vin{} \vin{} \vin{} Please, Mr.\ Flea!
\end{haiku}

\begin{haiku}
    {\FH 花の陰あかの他人はなかりけり}\hfill{\FH 一茶}

    \vin{} Beneath the cherry blossoms
    \vin{} \vin{} There are no
    \vin{} \vin{} \vin{} Strangers.
\end{haiku}

\begin{haiku}
    {\FH 鳴く猫に赤ん目をして手まりかな}\hfill{\FH 一茶}

    \vin{} The kitten is mewing;
    \vin{} \vin{} But the little girl is playing ball,
    \vin{} \vin{} \vin{} And only makes a face at him.
\end{haiku}

\begin{haiku}
    {\FH 鶯や御前へ出ても同じ聲}\hfill{\FH 一茶}

    \vin{} The nightingale!
    \vin{} \vin{} Even before His Lordship,
    \vin{} \vin{} \vin{} That same voice!
\end{haiku}

\begin{haiku}
    {\FH かたつぶり酒の肴に這はせけり}\hfill{\FH 其角}

    \vin{} Creep forward, O snail,
    \vin{} \vin{} That we may gaze on you
    \vin{} \vin{} \vin{} As relish to our wine!
\end{haiku}

\begin{haiku}
    {\FH 夕だちや家をめぐりて啼く家鴨}\hfill{\FH 其角}

    \vin{} A sudden summer shower;
    \vin{} \vin{} The ducks run round the house,
    \vin{} \vin{} \vin{} Quacking.
\end{haiku}

\begin{haiku}
    {\FH 黄菊白菊その外の名はなくもがな}\hfill{\FH 其角}

    \vin{} White chrysanthemums,
    \vin{} \vin{} Yellow chrysanthemums, ---
    \vin{} \vin{} \vin{} Would there were no other names!
\end{haiku}

\begin{haiku}
    {\FH うさくさをうしろに捨てて夏の月}\hfill{\FH 子規}

    \vin{} Behind me
    \vin{} \vin{} I cast all my care ---
    \vin{} \vin{} \vin{} The summer moon!
\end{haiku}

\begin{haiku}
    {\FH 工夫して花にランプを吊しけり}\hfill{\FH 子規}

    \vin{} Hanging a lantern on
    \vin{} \vin{} A blossoming bough ---
    \vin{} \vin{} \vin{} What pains I took!
\end{haiku}

\begin{haiku}
    {\FH 蜘蛛の巣に胡蝶のからのあはれかな}\hfill{\FH 子規}

    \vin{} In the spider's web
    \vin{} \vin{} Hang butterflies dead ---
    \vin{} \vin{} \vin{} A grievous sight!
\end{haiku}

\begin{haiku}
    {\FH 五月雨はただ降るものと覚えけり}\hfill{\FH 鬼貫}

    \vin{} I know well
    \vin{} \vin{} That the June rains
    \vin{} \vin{} \vin{} Just fall.
\end{haiku}

\begin{haiku}
    {\FH たふるればたふるるままの庭の草}\hfill{\FH 良寛}

    \vin{} The garden grasses:
    \vin{} \vin{} They fall,
    \vin{} \vin{} \vin{} And lie as they fall.
\end{haiku}

\begin{haiku}
    {\FH 盗人に取り残されし窓の月}\hfill{\FH 良寛}

    \vin{} The thief
    \vin{} \vin{} Left it behind ---
    \vin{} \vin{} \vin{} The moon at the window.
\end{haiku}

\begin{haiku}
    {\FH 鍋磨く音にまぎるる雨蛙}\hfill{\FH 良寛}

    \vin{} The sound
    \vin{} \vin{} Of the scouring of the saucepan
    \vin{} \vin{} \vin{} Blends with the green-frog's voice.
\end{haiku}

\begin{haiku}
    {\FH 雨の降る日はあはれなり良寛坊}\hfill{\FH 良寛}

    \vin{} On rainy days,
    \vin{} \vin{} The monk Ry\^okan
    \vin{} \vin{} \vin{} Feels sorry for himself.
\end{haiku}

\begin{haiku}
    {\FH 去年まで叱った瓜を手向けけり}\hfill{\FH 大江丸}

    \vin{} The melons, which last year
    \vin{} \vin{} I scolded him for eating,
    \vin{} \vin{} \vin{} I now offer to his spirit.
\end{haiku}

\begin{haiku}
    {\FH ふぐくはぬ人にはいはじ鰒の味}\hfill{\FH 大梅}

    \vin{} To a man who has not eaten
    \vin{} \vin{} A globe-fish,
    \vin{} \vin{} \vin{} We cannot speak of its flavour.
\end{haiku}

\begin{haiku}
    {\FH 名月や草木に劣る人の影}\hfill{\FH 梅室}

    \vin{} The full moon, and under the trees
    \vin{} \vin{} Their shadows --- how beautiful
    \vin{} \vin{} \vin{} Compared to mine!
\end{haiku}

\begin{haiku}
    {\FH 宿の春何もなきこそ何もあれ}\hfill{\FH 素堂}

    \vin{} In my hut this spring
    \vin{} \vin{} There is nothing ---
    \vin{} \vin{} \vin{} There is everything.
\end{haiku}

\begin{haiku}
    {\FH 廣き野をただ一のみや雉子の聲}\hfill{\FH 野明}

    \vin{} In one shrill cry
    \vin{} \vin{} The pheasant has swallowed
    \vin{} \vin{} \vin{} The broad field.
\end{haiku}

\begin{haiku}
    {\FH 御幸にも編笠ぬがぬ案山子かな}\hfill{\FH 団水}

    \vin{} Even before His Majesty,
    \vin{} \vin{} The scarecrow does not remove
    \vin{} \vin{} \vin{} Its plaited hat.
\end{haiku}

\begin{haiku}
    {\FH 月を松に懸けたり外しても見たり}\hfill{\FH 北枝}

    \vin{} I kept hanging the moon
    \vin{} \vin{} On the pine-tree and taking it off,
    \vin{} \vin{} \vin{} Gazing at it the while.
\end{haiku}

\begin{haiku}
    {\FH 流木やたぶりたぶりと春の川}\hfill{\FH 鳴雪}

    \vin{} A piece of wood ---
    \vin{} \vin{} Bobbity, bobbity it floats down
    \vin{} \vin{} \vin{} The spring river.
\end{haiku}

\begin{haiku}
    {\FH 犬を打つ石のさてなし冬の月}\hfill{\FH 太祇}

    \vin{} Not a single stone
    \vin{} \vin{} To throw at the dog, ---
    \vin{} \vin{} \vin{} wintry moon!
\end{haiku}

\begin{haiku}
    {\FH 蝶消えて魂我に返りけり}\hfill{\FH 和風}

    \vin{} The butterfly having disappeared,
    \vin{} \vin{} My spirit
    \vin{} \vin{} \vin{} Came back to me.
\end{haiku}

\begin{haiku}
    {\FH 陰に居て月に座敷をゆづりけり}\hfill{\FH 成美}

    \vin{} I sat in the shadows,
    \vin{} \vin{} And bequeathed the chamber
    \vin{} \vin{} \vin{} To the bright full moon.
\end{haiku}

\begin{haiku}
    {\FH 猫の子に嗅れてゐるや蝸牛}\hfill{\FH 才麿}

    \vin{} By a little kitten
    \vin{} \vin{} Sniffed at,
    \vin{} \vin{} \vin{} Creeps the slug unconcerned.
\end{haiku}

\begin{haiku}
    {\FH ながながと川ひとすぢや雪の原}\hfill{\FH 凡兆}

    \vin{} One long line of river
    \vin{} \vin{} Winds across
    \vin{} \vin{} \vin{} The snowy moor.
\end{haiku}
